---
title: "Wound chapter"
author: "Melissa Pavez-Fox"
date: "20/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(stringr)
library(magrittr)
library(readxl)
library(survival)
library(coxme)
library(effects)
library(lme4)
library(survminer)
library(ggplot2)
library(ggpubr)
library(ggeffects)
library(sjPlot)
library(visreg)
```

#Load datasets
```{r}
#Load wound data
wound_data = read.csv('C:/Users/mp660/OneDrive - University of Exeter/PhD/3rd chapter/Data for Markdown/wounds.csv')
#Load fem survival data
fem_data = read.csv('C:/Users/mp660/OneDrive - University of Exeter/PhD/3rd chapter/Data for Markdown/surv_fem.csv')
#Load male survival data
male_data = read.csv('C:/Users/mp660/OneDrive - University of Exeter/PhD/3rd chapter/Data for Markdown/surv_males.csv')
#Load survival data
surv_ages = read.csv('C:/Users/mp660/OneDrive - University of Exeter/PhD/3rd chapter/Data for Markdown/surv_ages.csv')

```

Format dataframes
```{r}
#REformat variables for analysis
surv_ages %<>% mutate(time_1 = as.numeric(time_1),
                       time_2 = as.numeric(time_2),
                       id = factor(id),
                       event = factor(event),
                       is_severe = factor(is_severe),
                       death = as.numeric(death),
                       group = factor(group),
                       group_size = as.numeric(group_size),
                       sex_ratio = as.numeric(round(sex_ratio, digits = 2)),
                       year_bim = factor(year_bim),
                       age = as.numeric(age),
                       sex = factor(sex),
                       is_mating = factor(is_mating),
                       severity = factor(severity, levels = c('1','0','2')))

fem_data %<>%   mutate(time_1 = as.numeric(time_1),
                       time_2 = as.numeric(time_2),
                       id = factor(id),
                       event = factor(event),
                       is_severe = factor(is_severe),
                       death = as.numeric(death),
                       group = factor(group),
                       group_size = as.numeric(group_size),
                       year_bim = factor(year_bim),
                       age = as.numeric(age),
                       sex = factor(sex),
                       is_mating = factor(is_mating),
                       n_kin1 = as.numeric(n_kin1),
                       n_kin2 = as.numeric(n_kin2),
                       rank = factor(rank),
                       severity = factor(severity, levels = c('1','0','2')))

male_data %<>%   mutate(time_1 = as.numeric(time_1),
                       time_2 = as.numeric(time_2),
                       id = factor(id),
                       event = factor(event),
                       is_severe = factor(is_severe),
                       death = as.numeric(death),
                       group = factor(group),
                       group_size = as.numeric(group_size),
                       year_bim = factor(year_bim),
                       age = as.numeric(age),
                       sex = factor(sex),
                       is_mating = factor(is_mating),
                       tenure = as.numeric(tenure),
                       severity = factor(severity, levels = c('1','0','2')))

#Create separate dataframes for logistic analyses (excluding NAs for rank and tenure)
male_rank = male_data %>% filter(!(is.na(tenure)))
fem_rank = fem_data %>% filter(!(is.na(rank)))

```


Q1: Does being injured increase the hazard of death?
```{r}
#######SURVIVAL ANALYSIS TO TEST IF WOUNDED ANIMALS HAVE INCREASED HAZARD COMPARED TO NON WOUNDED INDIVIDUALS####
#Find best model
model1 = coxme(Surv(time_1, time_2, death) ~ event + sex + is_mating + (1|id) + (1|year_bim), surv_ages)#best
model2 = coxme(Surv(time_1, time_2, death) ~ event*is_mating + sex + (1|id) + (1|year_bim), surv_ages)#NS
model3 = coxme(Surv(time_1, time_2, death) ~ event*sex + is_mating + (1|id) + (1|year_bim), surv_ages)#NS

#Model without random effect to test proportionality of hazards
no_rand_mod = coxph(Surv(time_1, time_2, death) ~ event + sex +  is_mating, surv_ages)
cox.zph(no_rand_mod)


###MODEL TO TEST IF SEVERE INJURIES HAVE A HIGHER HAZARD THAN NON-SEVERE INJURIES
#Find if interactions are significant
model_sev1 = coxme(Surv(time_1, time_2, death) ~ severity*is_mating + sex + (1|id) + (1|year_bim), surv_ages)#NS
model_sev2 = coxme(Surv(time_1, time_2, death) ~ severity*sex + is_mating + (1|id) + (1|year_bim), surv_ages)#Interaction significant
model_sev3 = coxme(Surv(time_1, time_2, death) ~ severity + sex + is_mating + (1|id) + (1|year_bim), surv_ages)#NS

#Check for PH
check = coxph(Surv(time_1, time_2, death) ~ severity*sex + is_mating, surv_ages)
cox.zph(check)#No issues

rm(model1,model2,model3,model_sev1,model_sev2,model_sev3)
```

Q2: Does sociality influence the probability of being injured?

#Social status
```{r}
#-#-#-FEMALES#-#-#-
#ALL INJURIES##
logit_fem1 = glmer(event ~ rank + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_rank, family = binomial)#NS
logit_fem2 = glmer(event ~ rank*scale(age_bimon) + is_mating +  (1|id) + (1|year_bim), data = fem_rank, family = binomial)#SIGNIFICANT
logit_fem3 = glmer(event ~ rank*is_mating + scale(age_bimon) +  (1|id) + (1|year_bim), data = fem_rank, family = binomial)#NS


#Severe injuries
logit_sev_F1 = glmer(is_severe ~ rank + is_mating + scale(age_bimon)  + (1|id) +  (1|year_bim), data = fem_rank, family = binomial)
logit_sev_F2 = glmer(is_severe ~ rank*is_mating + scale(age_bimon)  + (1|id) + (1|year_bim), data = fem_rank, family = binomial)#NS
logit_sev_F3 = glmer(is_severe ~ rank*scale(age_bimon) + is_mating  + (1|id) + (1|year_bim), data = fem_rank, family = binomial)#NS


#-#-#-#-MALES#-##--
logit_male1 = glmer(event ~ scale(tenure) + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = male_rank, family = binomial)
logit_male2 = glmer(event ~ scale(tenure)*is_mating + scale(age_bimon)  + (1|id) + (1|year_bim), data = male_rank, family = binomial)#NS
logit_male3 = glmer(event ~ scale(tenure)*scale(age_bimon)  + is_mating + (1|id)  + (1|year_bim), data = male_rank, family = binomial)#SIGNIFICANT

#Check collinearity between tenure and age
car::vif(logit_male1)#no issues vif = 1.01

#Severe injuries
logit_sev_M1 = glmer(is_severe ~ scale(tenure) + scale(age_bimon) + is_mating +  (1|id) + (1|year_bim), data = male_rank, family = binomial)#
logit_sev_M2 = glmer(is_severe ~ scale(tenure)*is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = male_rank, family = binomial)#NS
logit_sev_M3 = glmer(is_severe ~ scale(tenure)*scale(age_bimon)  + is_mating + (1|id) + (1|year_bim) , data = male_rank, family = binomial)#SIGNIFICANT



rm(logit_fem1,logit_fem2,logit_fem3,logit_sev_F1,logit_sev_F2,logit_sev_F3,
  logit_male1,logit_male2,logit_male3,logit_sev_M1,logit_sev_M2,logit_sev_M3)
```


#Affiliative partners
```{r}
#-#-#-CLOSE KIN#-#-#-
#ALL INJURIES##
logit_fem1 = glmer(event ~ scale(n_kin2) + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_data, family = binomial)
logit_fem2 = glmer(event ~ scale(n_kin2)*scale(age_bimon) + is_mating  +  (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS
logit_fem3 = glmer(event ~ scale(n_kin2)*is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS


#Severe injuries
logit_sev_F1 = glmer(is_severe ~ scale(n_kin2) + is_mating + scale(age_bimon)  + (1|id) +  (1|year_bim), data = fem_data, family = binomial)
logit_sev_F2 = glmer(is_severe ~ scale(n_kin2)*is_mating + scale(age_bimon) +  (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS
logit_sev_F3 = glmer(is_severe ~ scale(n_kin2)*scale(age_bimon) + is_mating + (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS



#-#-#-EXTENDED KIN#-#-#-
#ALL INJURIES##
logit_fem1 = glmer(event ~ scale(n_kin1) + is_mating + scale(age_bimon)  + (1|id) + (1|year_bim), data = fem_data, family = binomial)#
logit_fem2 = glmer(event ~ scale(n_kin1)*scale(age_bimon) + is_mating  +  (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS
logit_fem3 = glmer(event ~ scale(n_kin1)*is_mating + scale(age_bimon)  + (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS


#Severe injuries
logit_sev_F1 = glmer(is_severe ~ scale(n_kin1) + is_mating + scale(age_bimon)  + (1|id) +  (1|year_bim), data = fem_data, family = binomial)
logit_sev_F2 = glmer(is_severe ~ scale(n_kin1)*is_mating + scale(age_bimon)  + (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS
logit_sev_F3 = glmer(is_severe ~ scale(n_kin1)*scale(age_bimon) + is_mating  + (1|id) + (1|year_bim), data = fem_data, family = binomial)#NS



rm(logit_fem1,logit_fem2,logit_fem3,logit_sev_F1,logit_sev_F2,logit_sev_F3)
```


#Group size
```{r}

###ALL INJURIES###
logit_gsize = glmer(event ~ scale(group_size) + is_mating + scale(age_bimon) + sex + (1|id) + (1|year_bim), data = surv_ages, family = binomial)
logit_gsizeI1 = glmer(event ~ scale(group_size)*sex + scale(age_bimon) + is_mating + (1|id) + (1|year_bim), data = surv_ages, family = binomial)#significant
logit_gsizeI2 = glmer(event ~ scale(group_size)*scale(age_bimon) + is_mating + sex + (1|id) + (1|year_bim) , data = surv_ages, family = binomial)#NS
logit_gsizeI3 = glmer(event ~ scale(group_size)*is_mating + scale(age_bimon) + sex + (1|id) + (1|year_bim) , data = surv_ages, family = binomial)#significant



###SEVERE INJURIES#-#-#
logit_gsize_sev = glmer(is_severe ~ scale(group_size) + is_mating + scale(age_bimon) + sex + (1|id) + (1|year_bim), data = surv_ages, family = binomial)
logit_gsizeI1_sev = glmer(is_severe ~ scale(group_size)*sex + scale(age_bimon) + is_mating + (1|id) + (1|year_bim), data = surv_ages, family = binomial)#significant
logit_gsizeI2_sev = glmer(is_severe ~ scale(group_size)*scale(age_bimon) + is_mating + sex + (1|id) + (1|year_bim) , data = surv_ages, family = binomial)#NS
logit_gsizeI3_sev = glmer(is_severe ~ scale(group_size)*is_mating + scale(age_bimon) + sex + (1|id) + (1|year_bim), data = surv_ages, family = binomial)#significant


rm(logit_gsize,logit_gsizeI1,logit_gsizeI2,logit_gsizeI3, logit_gsize_sev,logit_gsizeI1_sev,logit_gsizeI2_sev,logit_gsizeI3_sev )
```


Q3: Does sociality affect the hazard of death after being injured

Survival models to test effect of social status on injured individuals
```{r}
#######SURVIVAL ANALYSIS FOR FEMALES#-#-#-#-
#Model for all injuries
fem = coxme(Surv(time_1, time_2, death) ~ event*rank  + is_mating + group_size +  (1|id) + (1|year_bim), fem_data)#

#CHECK PH
check = coxph(Surv(time_1, time_2, death) ~ event*rank  + is_mating + group_size, fem_data)
cox.zph(check)#NO issues

#SEVERITY MODELS FOR FEMALES
fem_sev = coxme(Surv(time_1, time_2, death) ~ is_severe*rank + group_size + is_mating + (1|id) + (1|year_bim), fem_data)#

#Test proportionality of hazards
check = coxph(Surv(time_1, time_2, death) ~ is_severe*rank  +  group_size + is_mating, fem_data)#No issues with proportionality of hazards
cox.zph(check)#No issues

####SURVIVAL ANALYSIS FOR MALES#-#-#-#-#-
#Model for all injuries
male = coxme(Surv(time_1, time_2, death) ~ event*tenure  + group_size + is_mating + (1|id) + (1|year_bim), male_data)#

model_check = coxph(Surv(time_1, time_2, death) ~ event*tenure + is_mating + group_size , male_data)#
cox.zph(model_check)#No issues

#SEVERITY MODELS FOR MALES
male_sev = coxme(Surv(time_1, time_2, death) ~ is_severe*tenure  + group_size + is_mating + (1|id) + (1|year_bim), male_data)#NS

#Test proportionality of hazards
check = coxph(Surv(time_1, time_2, death) ~ is_severe*tenure  + group_size + is_mating, male_data)
cox.zph(check)#No issues

###
###IMPUTATION OF MISSING DATA#####
library(mice)
imp = mice(male_data, defaultMethod = 'pmm', m = 20)#20 iterations using predictive mean matching

coximpute = with(imp,coxph(Surv(time_1, time_2, death) ~ event*tenure +  group_size + is_mating, male_data))
summary(pool(coximpute))

cox_sevimpute = with(imp,coxph(Surv(time_1, time_2, death) ~ is_severe*tenure +  group_size + is_mating, male_data))
summary(pool(cox_sevimpute))


rm(model_check, fem,check, male, male_sev,fem_sev,imp,coximpute,cox_sevimpute,for_plot)
   
```


Survival models to test the effect of affiliative partners on injured individuals
```{r}
#EXTENDED KIN
#ALL INJURIES####
fem = coxme(Surv(time_1, time_2, death) ~ event*n_kin1  +  group_size + is_mating + (1|id) + (1|year_bim), fem_data)#

#CHeck PH 
check = coxph(Surv(time_1, time_2, death) ~ event*n_kin1 + group_size + is_mating, fem_data)
cox.zph(check)#no issues

#SEVERE INJURIES###-#-
#MODEL
fem_sev = coxme(Surv(time_1, time_2, death) ~ is_severe*n_kin1  + group_size + is_mating + (1|id) + (1|year_bim), fem_data)#

#CHECK PH model
check = coxph(Surv(time_1, time_2, death) ~ is_severe*n_kin1  + group_size + is_mating, fem_data)
cox.zph(check)#No issues

#MODELS FOR CLOSE KIN#-#-#-#-
#model
fem = coxme(Surv(time_1, time_2, death) ~ event*n_kin2  + group_size + is_mating + (1|id) + (1|year_bim), fem_data)#
#Chekc PH
check = coxph(Surv(time_1, time_2, death) ~ event*n_kin2   + group_size + is_mating, fem_data)
cox.zph(check)#No issues

#SEVERE INJURIES###-#-
#Model
fem_sev = coxme(Surv(time_1, time_2, death) ~ is_severe*n_kin2  + group_size + is_mating + (1|id) + (1|year_bim), fem_data)#

#Check PH
check = coxph(Surv(time_1, time_2, death) ~ is_severe*n_kin2  + group_size + is_mating, fem_data)
cox.zph(check)#No issues


rm(check, fem,fem_sev)      
```


Post hoc mediation analysis
```{r}
#Iacobucci method (only works for logit models without interactions)
#Not useful for survival models, logit models used instead
fem_mediation = fem_data %>% filter(!(status == 2))#remove culled individuals

#Equation 1: Survival ~ affiliative partners + covariates
Y = glmer(death ~ n_kin1 + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_mediation, family = binomial)

#Equation 2: Mediator ~ affiliative partners + covariates
M = glmer(event ~ n_kin1 + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_mediation, family = binomial)

#Equation 3: mediation effect of injuries on the relationship between sociality and survival
Y2 = glmer(death ~ n_kin1 + event + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_mediation, family = binomial)


#Iacobucci method
#get estimate and SE for affiliative partners from Eq.2
par_a = as.numeric(fixef(M)[2])#parameter a
se_a = as.numeric(summary(M)$coef[2, 2, drop = FALSE])#standard error a

#get estimate and se for injuries from Eq. 3
par_b = as.numeric(fixef(Y2)[3])
se_b = as.numeric(summary(Y2)$coef[3, 2, drop = FALSE])

#compute standardised elements
Za = par_a/se_a
Zb = par_b/se_b
#product:
Zab = Za*Zb
#collective standard error
coll_se = sqrt((Za^2) + (Zb^2) + 1)

#test significance of mediation
mediation = Zab/coll_se#If > |1.96| significant

#Direct effect of social status
fem_surv = fem_rank %>% filter(!(status == 2))#remove culled individuals
male_surv = male_rank %>% filter(!(status == 2))#remove culled individuals
#Females
fem = glmer(death ~ rank + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_surv, family = binomial)
male = glmer(death ~ scale(tenure) + is_mating + scale(age_bimon) + (1|id), data = male_surv, family = binomial)#year_bim removed for convergence issues

#Direct effect of group size
group_data = surv_ages %>% filter(!(status == 2))#remove culled individuals
group = glmer(death ~ scale(group_size) + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = group_data, family = binomial)


```


#PLOTS
Figure 2: effect of injuries on survival 
```{r}
#Figure 2A: survival curves for injured vs uninjured animals
#rerun model without random effects (can not be included for using survfit)
cox_model = coxph(Surv(time_1, time_2, death) ~ event + sex + is_mating, surv_ages)

#Survival curves: plot is for females during mating season, but is similar for males
cox_plot = survfit(cox_model, newdata = data.frame(sex = "F", is_mating = "1", event = c("0","1")))
ggsurvplot(cox_plot, data = surv_ages,
           conf.int = TRUE, size = 1,
           conf.int.style = "ribbon",
           conf.int.alpha = 0.1,
           size = 1,
           censor.size=0.1, 
           linetype = c("dashed","solid"),
           lengend = "bottom",
           legend.title = "Injury status",
           legend.labs = c("Uninjured","Injured"),
           break.time.by = 5,
           xlab = "Age (years)",
           xlim = c(4,30),
           palette = c("grey47","indianred"))

#Figure 2B: hazard ratio based on severity by sex
#Change factor order of severity to leave uninjured as intercept
surv_ages1 = surv_ages %<>% mutate(severity = factor(severity, levels = c('0','1','2')))
#Run model
cox_model = coxph(Surv(time_1, time_2, death) ~ severity*sex + is_mating, surv_ages1)

#Using alternative packages

plot_surv = plot_model(cox_model, type = "int", ci.lvl = 0.95, ci.style = "bar", dot.size = 3.5, line.size = 1,
                       colors = c("lightcoral","darkslategray4"))
plot_surv + theme_bw() 

```


Figure 3: Incidence of injuries as function of social status in females and males
```{r}

#FIGURE 3A: Social status*Age Females
fit= glmer(event ~ rank*scale(age_bimon) + is_mating + (1|id) + (1|year_bim), data = fem_rank, family = binomial)

#Get fitted values
ef1 <- effect(term="rank*scale(age_bimon)", 
              xlevels=list(rank = c("High","Low"),
                           age_bimon = seq(4,29,1)),
              mod=fit)

#Convert to dataframe
efdata1 <-as.data.frame(ef1)
efdata1$rank = as.factor(efdata1$rank)#factor
fem_rank$event1 = as.integer(as.character(fem_rank$event))#integer
fem_rank %<>% mutate(event1 = ifelse(event1 == 1, 0.18, 0))#Match event data point to y-axis length

#Plot
Fig3A = ggplot(efdata1, aes(x=age_bimon, y=fit, color=rank,group=rank)) + 
  geom_line(size=1.2, aes(linetype = rank)) +
  geom_ribbon(aes(ymin=fit-se, ymax=fit+se, fill= rank),alpha=0.3) + 
   #Add raw data
  geom_jitter(data = fem_rank,
              aes(y = event1), colour = "antiquewhite4",
              size = 4, alpha = 0.2, height = 0.0015) +
  scale_color_manual(values = c("pink4","goldenrod1")) +
  scale_fill_manual(values = c("pink4","goldenrod1")) +
  scale_y_continuous(expand = c(0, 0.01)) +
  scale_x_continuous(breaks=seq(0,29,5)) +
  labs(x= "Age (years)", y="Probability of being injured", 
       color="Social status", fill="Social status", linetype = "Social status") + theme_classic() + theme(text=element_text(size=20), legend.position = c(0.2,0.75),legend.text = element_text(size = 18))


#FIGURE 3B: Social status*Age Males
fit = glmer(event ~ scale(tenure)*scale(age_bimon) + is_mating + (1|id) + (1|year_bim), data = male_rank, family = binomial)

#Get 20th and 80th quantiles for tenure
x = quantile(male_rank$tenure, probs = c(0.2,0.8))

#Get fitted values
ef1 <- effect(term ="scale(tenure)*scale(age_bimon)",
              xlevels = list(tenure = c(x[1],x[2]),
                            age_bimon = seq(4,26,1)),               
                            mod=fit)

#Convert to dataframe
efdata1 <- as.data.frame(ef1)
efdata1$tenure = as.factor(as.character(efdata1$tenure))#factor
male_rank$event1 = as.integer(as.character(male_rank$event))#integer
male_rank %<>% mutate(event1 = ifelse(event1 == 1, 0.085, 0))#Match datapoints to y-axis length

#Plot
Fig3B = ggplot(efdata1, aes(x=age_bimon, y=fit, color= tenure, group= tenure)) + 
  geom_line(size=1.2, aes(linetype = tenure)) +
  geom_ribbon(aes(ymin= fit-se, ymax= fit+se, fill= tenure),alpha=0.3) + 
   #Add raw data
  geom_jitter(data = male_rank,
              aes(y = event1), colour = "antiquewhite4",
              size = 4, alpha = 0.2, height = 0.0015)  +
  scale_color_manual(labels = c("High","Low"),values = c("pink4","goldenrod1")) +
  scale_fill_manual(labels = c("High","Low"), values = c("pink4","goldenrod1")) +
  scale_x_continuous(breaks=seq(0,26,5)) +
  scale_y_continuous(breaks=seq(0,0.1,0.025)) +
  labs(x= "Age (years)", y="Probability of being injured", 
       color="Social status", fill="Social status", linetype = "Social status") + theme_classic() + 
       theme(text=element_text(size=20), legend.position = c(0.2,0.75), 
             legend.text =  element_text(size = 18))
                                                                  

```


Figure 4: incidence of injuries as function of affiliative partners 
```{r}

#FIGURE 4: Affiliative partners
fit = glmer(event ~ scale(n_kin1) + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_data, family = binomial)

#Get fitted values
ef1 <- effect(term="scale(n_kin1)",
              xlevels=list(n_kin1 = 0:38),
              mod=fit)
#Convert to dataframe
efdata1 <- as.data.frame(ef1)
fem_data$event1 = as.integer(as.character(fem_data$event))#integer
fem_data %<>% mutate(event1 = ifelse(event1 == 1, 0.022, 0))#Match datapoints to y-axis length

#Plot
Fig4A = ggplot(efdata1, aes(x= n_kin1, y= fit)) + 
  geom_line(size=1.2, colour = "slategray") +
  geom_ribbon(aes(ymin= fit-se, ymax= fit+se),alpha=0.3, fill = "slategray") + 
  #Add raw data
  geom_jitter(data = fem_data,
              aes(y = event1), colour = "antiquewhite4",
              size = 4, alpha = 0.2, height = 0.0002)  +
  scale_x_continuous(breaks=seq(0,38,10)) +
  scale_y_continuous(breaks=seq(0,0.025,0.01)) +
  labs(x= "Affiliative partners in the group", y="Probability of being injured") + 
  theme_classic() + theme(text=element_text(size=20), legend.position = c(0.2,0.75), 
                          legend.text =  element_text(size = 18))

```

#SUPLEMENTARY ANALYSES

Healing time and Sampling effort
```{r}
#COMPUTE HEALING TIME##-#-#-#-#-#
#Create a healing dataframe
healing = wound_data
#Keep only outcomes related to recovery
healing %<>% filter(Outcome %in% c('Healed'))
#Disregard individuals with unknown date of released
healing %<>% filter(Date_Released != "")#only 250 cases

#Compute time to heal 
healing$healing_time <- difftime(healing$Date_Released,healing$Sick_date, units = "days")

#Check how many individuals have time to heal == 0
healing$healing_time = as.integer(healing$healing_time)
sum(healing$healing_time == 0)#13

#Remove animals with healing == 0, some are old wounds, some are incorrect follow ups, and others are probably wrong
healing %<>% filter(healing_time != '0')#261 cases

#Exclude old wounds as the time to heal for those would be longer than the computed time
healing %<>% filter(is_old == '0')#250

#Check if there are individuals with date of released = sick date
sum(healing$Sick_date == healing$Date_Released)#NOne

#Disregard follow ups and limping
healing %<>% filter(IsFollowUp == '0',Illness %in% c('abrasion','wound(s)'))

#Visualise distribution of healing time
hist(healing$healing_time)#Two healing times above 200 days (probably incorrect)

#Remove unreliable healing times
healing %<>% filter(healing_time <= 200)

#COMPUTE MEAN AND SD
mean(healing$healing_time)
sd(healing$healing_time)

#SEVERE INJURIES#
healing %>% group_by(is_severe) %>% summarise(mean(healing_time),sd(healing_time))

rm(healing)
#######-#-#-#-#-#-#-#--

#SAMPLING EFFORT #-#-#-#-##-
#COMPUTE AVERAGE TIME OF UPDATE FOR RECORDS
health_update_wound <- as.data.frame(unique(wound_data$Sick_date))
health_update_wound %<>% rename(Date = `unique(wound_data$Sick_date)`) %>% filter(Date != "")
health_update_wound$Date = as.Date(health_update_wound$Date, format = '%Y-%m-%d')
health_update_wound$Date = format(health_update_wound$Date, format = '%d-%b-%Y')
health_update_released <- as.data.frame(unique(wound_data$Date_Released))
health_update_released %<>% rename(Date = `unique(wound_data$Date_Released)`) %>% filter(Date != "")
#Unify date format
for(i in 1: nrow(health_update_released)){ 
  splitdate=str_split(health_update_released$Date[i],"-")
  year=ifelse(nchar(splitdate[[1]][3])==4,splitdate[[1]][3],paste("20",splitdate[[1]][3], sep=""))
  health_update_released$Date[i]= paste(splitdate[[1]][1],splitdate[[1]][2],year, sep="-") 
}

health_update <- rbind(health_update_released,health_update_wound)
health_update$Date = as.Date(health_update$Date, format = "%d-%b-%Y")
health_update = as.data.frame(unique(health_update$Date))
health_update %<>% rename(Date = `unique(health_update$Date)`)
health_update = health_update[order(as.Date(health_update$Date, format="%d-%m-%Y")),]
health_update = as.data.frame(health_update)
health_update %<>% rename(Date = health_update)
health_update %<>% filter(!(is.na(Date)))#remove NA

#COMPUTE AVERAGE TIME OF UPDATING INJURIES
mean(diff(health_update$Date))#5.6 days

#COMPUTE TIME OF UPDATING PER YEAR
for(i in 1: nrow(health_update)){ 
  splitdate=str_split(health_update$Date[i],"-")
  year=ifelse(nchar(splitdate[[1]][1])==4,splitdate[[1]][1],paste("20",splitdate[[1]][3], sep=""))
  health_update$year[i]= year 
}

#Mean update per year
health_update %>% group_by(year) %>% summarise(mean(diff(Date)))

#Remove unnecessary info
rm(health_update,health_update_released,health_update_sick,splitdate)

#COMPUTE UPDATES FOR EACH INDIVIDUAL#-#-#-#-#-#-#-#
check = wound_data %>% arrange(ANIMAL_ID,Sick_date)
check$Sick_date = as.Date(check$Sick_date, format = '%Y-%m-%d')
check$Date_Released = as.Date(check$Date_Released, format = '%Y-%m-%d')
check$timetoupdate= rep(NA,nrow(check))

num.zeros = 0
num.followup = 0
num.duplicates = 0

for(i in 1:nrow(check)){
  
  if (i != 1){
    if(check$ANIMAL_ID[i] == check$ANIMAL_ID[i -1]){
      if(check$Sick_date[i] == check$Sick_date[i-1]){
        check$timetoupdate[i] = NA
        num.duplicates = num.duplicates + 1#compute duplicates
      }
      else {
        if(check$IsFollowUp[i] == '1'){
          check$timetoupdate[i] = check$Sick_date[i] - check$Sick_date[i-1]
          num.followup = num.followup +1 #compute number of follow ups
        }
        else {
          check$timetoupdate[i] = check$Date_Released[i] - check$Sick_date[i]
        }
      }
    }
    else {
      check$timetoupdate[i] = check$Date_Released[i] - check$Sick_date[i]
    }
  }
  if(!is.na(check$timetoupdate[i])){
    if(check$timetoupdate[i] == "0"){
      check$timetoupdate[i] = NA
      num.zeros = num.zeros +1#compute number of times date released = sick date
    }
  }
}


#Compute average time of update for individuals
check %<>% filter(!is.na(timetoupdate)) 
mean(check$timetoupdate)#42.17 days
#Compute time to update per year
check %>% group_by(year) %>% summarise(mean(timetoupdate))

rm(check,health_data)

```


 LOGISTIC MODELS RATHER THAN COX MODELS
Logistic regression to test the effect of injuries on survival (Q1)
```{r}
#Exclude last bimonth for individuals that were removed [status = 2]
for_logit = surv_ages %>% filter(!(status == 2))

#Model to test if being injured is associated to death within the next to months
logit = glmer(death ~ event + is_mating + scale(age_bimon) + sex + (1|id) + (1|year_bim), data = for_logit, family = binomial)

#Model to test if severe injuries are associated to death
logit_sev = glmer(death ~ severity*sex + is_mating + (1|id) + (1|year_bim), data = for_logit, family = binomial)#significant; term removed due to convergence issues

```

Logistic regression to test the buffering effect of social status (Q3a)
```{r}
#Remove last bimonth for individuals that were removed
fem_data1 %<>% filter(!(status == 2))
male_data1 %<>% filter(!(status == 2))

#FEMALES#
#All injuries###
logit_rankF = glmer(death ~ event*rank + is_mating + scale(age_bimon) + scale(group_size) + (1|id) + (1|year_bim), data = fem_data1, family = binomial)

#Severe injuries##
logit_rankF_sev = glmer(death ~ is_severe*rank + is_mating + scale(age_bimon) + scale(group_size) + (1|id) + (1|year_bim), data = fem_data1, family = binomial)#

#MALES#
#All injuries###
logit_rankM = glmer(death ~ event*scale(tenure) + is_mating + scale(age_bimon) + scale(group_size) + (1|id) + (1|year_bim), data = male_data1, family = binomial)

#Severe injuries#
logit_rankM_sev = glmer(death ~ is_severe*scale(tenure) + is_mating +  (1|id) + (1|year_bim), data = male_data1, family = binomial)#terms removed due to convergence issues

```

Logistic regression to test the buffering effect of social capital (Q3b)
```{r}
#FEMALES-extended kin
logit_kin1 = glmer(death ~ event*n_kin1 + is_mating + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_data1, family = binomial)#NS(term removed due to convergence issues)
#Severe injuries
logit_kin1_sev = glmer(death ~ is_severe*n_kin1 + is_mating + scale(age_bimon) + scale(group_size) + (1|id) + (1|year_bim), data = fem_data1, family = binomial)#

#Close kin
logit_kin2 = glmer(death ~ event*n_kin2 + is_mating + scale(age_bimon) + scale(group_size) + (1|id) + (1|year_bim), data = fem_data1, family = binomial)#NS
#Severe injuries
logit_kin2_sev = glmer(death ~ is_severe*n_kin2 + scale(age_bimon) + (1|id) + (1|year_bim), data = fem_data1, family = binomial)#(term removed due to convergence issues)

```


SUPPLEMENTARY FIGURES
Figure S1: Samples size and death events for injured and not injured animals
```{r}
#Load dataframe for plot
unique_ids = read.csv('C:/Users/mp660/OneDrive - University of Exeter/PhD/3rd chapter/Data for Markdown/unique_ids.csv')
#Subset injured and not injured animals
injured = unique_ids %>% filter(is_injured == '1')
not_injured = unique_ids %>% filter(is_injured == '0')

#Summarise data for plot
for_plot_inj = injured %>% group_by(sex) %>% summarise(deaths = sum(!is.na(DOD)), rem = sum(!is.na(removal)),
                                                       alive = sum(is.na(DOD) & is.na(removal))) %>%
                                             mutate(status = 'injured',
                                                    sex = ifelse(sex == 'F', 'Female','Male'))
for_plot_NI = not_injured %>% group_by(sex) %>% summarise(deaths = sum(!is.na(DOD)), rem = sum(!is.na(removal)),
                                                       alive = sum(is.na(DOD) & is.na(removal))) %>%
                                                mutate(status = 'not injured',
                                                       sex = ifelse(sex == 'F', 'Female','Male'))

#Change to long format
library(reshape2)
inj = for_plot_inj %>% select(sex,deaths,rem,alive)
inj = melt(inj, id=c('sex'))

ninj = for_plot_NI %>% select(sex,deaths,rem,alive)
ninj = melt(ninj, id=c('sex'))

#Plots
inj %>% ggplot(aes(x = variable, y = value, fill = sex)) +
        geom_bar(stat="identity", alpha= 0.75, position=position_dodge(), colour = "black", width = 0.65,
                 aes(linetype = sex)) +
        labs(x= "", y= "Number of injured individuals",fill= "Sex", linetype = "Sex") +
        scale_fill_manual(values= c("indianred1","darkslategray4"))+ 
        coord_flip() +
        theme_bw()

ninj %>% ggplot(aes(x = variable, y = value, fill = sex)) +
        geom_bar(stat="identity", alpha= 0.85, position=position_dodge(), colour = "black", width = 0.65,
                 aes(linetype = sex)) +
        labs(x= "", y= "Number of uninjured individuals",fill= "Sex", linetype = "Sex") +
        scale_fill_manual(values= c("indianred1","darkslategray4"))+ 
        coord_flip() +
        theme_bw()


```


Fig S2: Incidence of severe injuries as function of social status 
```{r}

#FigS2 Probability of being severely injured as function of social status in Males
#Social status males (females not significant)
fit = glmer(is_severe ~ scale(tenure)*scale(age_bimon)  + is_mating + (1|id) + (1|year_bim) , data = male_rank, family = binomial)
#Get 20th and 80th quantiles for tenure
x = quantile(male_rank$tenure, probs = c(0.2,0.8))

#Get fitted values
ef1 <- effect(term ="scale(tenure)*scale(age_bimon)",
              xlevels = list(tenure = c(x[1],x[2]),
                            age_bimon = seq(4,26,1)),               
                            mod=fit)

#Convert to dataframe
efdata1 <- as.data.frame(ef1)
efdata1$tenure = as.factor(as.character(efdata1$tenure))#factor
male_rank$event1 = as.integer(as.character(male_rank$is_severe))#integer
male_rank %<>% mutate(event1 = ifelse(event1 == 1, 0.07, 0))#Match datapoints to y-axis length

#Plot
FigS2 = ggplot(efdata1, aes(x=age_bimon, y=fit, color= tenure, group= tenure)) + 
   geom_line(size=1.2, aes(linetype = tenure)) +
  geom_ribbon(aes(ymin= fit-se, ymax= fit+se, fill= tenure),alpha=0.3) + 
   #Add raw data
  geom_jitter(data = male_rank,
              aes(y = event1), colour = "antiquewhite4",
              size = 4, alpha = 0.2, height = 0.0007)  +
  scale_color_manual(labels = c("High","Low"),values = c("pink4","goldenrod1")) +
  scale_fill_manual(labels = c("High","Low"), values = c("pink4","goldenrod1")) +
  scale_x_continuous(breaks=seq(0,26,5)) +
  scale_y_continuous(breaks=seq(0,0.8,0.02)) +
  labs(x= "Age (years)", y="Probability of being severely injured", 
       color="Social status", fill="Social status") + theme_classic() + 
       theme(text=element_text(size=20), legend.position = c(0.2,0.75), 
             legend.text =  element_text(size = 18))

```

